---
title: "`r params$title %||% 'Metabolomics Data Quality Report'`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 2
    toc_float: true
params:
  title:
    value: "Metabolomics Data Quality Report"
  notes:
    value: ""
  plots:
    value: !r character(0)
  p:
    value: !r NULL
  d:
    value: !r NULL
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(dev = "svg")
library(htmltools)
library(dplyr)
library(QCcorrection)
```
# Raw Data
```{r}
QCcorrection:::report_text_data_req()
QCcorrection:::report_text_withheld_columns(p,d)
```

### Raw Data Summary and Structural Checks

**Raw data summary and structural checks.** 
```{r}
QCcorrection:::report_text_data_inspection()
QCcorrection:::ui_basic_info(params$d$cleaned$df, params$d$cleaned$replacement_counts, params$d$cleaned$non_numeric_cols, params$d$cleaned$duplicate_mets, params$d$cleaned$blank_df, params$d$cleaned$below_blank_threshold)
```

### Filter Missing Values in Raw Data
```{r}
QCcorrection:::report_text_mv_filter()
QCcorrection:::ui_filter_info(params$d$filtered$mv_removed_cols, params$d$filtered$mv_cutoff, params$d$filtered$qc_missing_mets)
```

### Raw Data Metabolite Correlations
This sections shows pairs of metabolites that have Pearson's r correlations within the specified range. To investigate linear relationships between metabolites, Pearson's r is computed for each pair. A strong positive linear correlation (Pearson's r near 1) means that as one metabolite increases, the other metabolite consistently increases proportionally.

All pairwise correlations are computed, but we only allow pairs with a strong positive linear correlations to be displayed here.

```{r}
QCcorrection:::report_text_correlations()
if (!is.null(params$d$raw_corr)) {
  QCcorrection:::ui_corr_range_info(params$d$raw_corr, params$p$raw_corr_threshold)
} else {
  htmltools::tags$span(style = "font-weight:bold;", "Metabolite correlations were not computed.")
}
```

Two metababolites might have a strong positive linear correlation without a biological explanation if they are from the same chromotography peak, but viewed under different filters or
they have similar signal drift patterns that confounds biological signal.
If a pair of metabolites has a strong positive linear correlation without a biological explanation, further investigation is needed to verify they are not the same compound.

# Imputation, Signal Drift Correction, and Transformation.

### Missing Value Imputation
```{r}
QCcorrection:::report_text_imputation(params$p, params$d)
```

### Correction Description
```{r}
QCcorrection:::report_text_correction(params$p, params$d)
```

### Post-correction Filtering
#### QC RSD Fitlering

```{r}
QCcorrection:::report_text_rsd_cal()
QCcorrection:::ui_postcor_filter_info(params$d$filtered_corrected, params$p$remove_imputed, params$p$rsd_cutoff, params$p$post_cor_filter)
```

### Transformation Description
```{r}
QCcorrection:::report_text_transformation(params$p, params$d)
```

### Candidate Extreme Values

```{r}
QCcorrection:::report_text_ev_detection()
QCcorrection:::ui_outliers(params$p, params$d) 
```

### Transformed and Corrected Metabolite Correlations
```{r}
if (!is.null(params$d$tc_corr)) {
  QCcorrection:::ui_corr_range_info(params$d$tc_corr, params$p$tc_corr_threshold)
} else {
  htmltools::tags$span(style = "font-weight:bold;", "Metabolite correlations were not computed.")
}
```
# Visualization

### Visualize the Correction with Metabolite Scatter Plots

```{r, echo=FALSE, fig.show='hold', out.width='49%', fig.width=5, fig.height=5}
print(params$plots[["Metabolite Scatter 1"]])
print(params$plots[["Metabolite Scatter 2"]])
```

```{r}
QCcorrection:::report_text_scatter_intro(params$p, params$d)
```

### RSD Comparison

```{r, echo=FALSE, fig.width=7.5, fig.height=4.5}
print(params$plots[["RSD Comparison"]])
```

To judge how well the correction method worked, we can visualize the change in variation with the relative standard deviation (RSD) comparison plots. RSD is sometimes called coefficient of variation (CV) and is computed as RSD = (standard deviation / mean) * 100%. High RSD indicates that the values for that metabolite are more spread out around the mean, which may be an expected result in the experimental samples. For QC samples, RSD should decrease after correction.
```{r}
QCcorrection:::report_text_rsd_intro(params$p, params$d)
```

```{r}
QCcorrection:::ui_rsd_stats(params$p, params$d)
```

### PCA Comparison

```{r, echo=FALSE, fig.width=8.333, fig.height=4.417}
print(params$plots[["PCA Comparison"]])
```

```{r}
QCcorrection:::report_text_pca_intro(params$p, params$d)
```

```{r, echo=FALSE, fig.width=8.75, fig.height=4.417}
print(params$plots[["PCA Loading"]])
```

The loading values for the top 10 metabolites for each PC are plotted above. The magnitude of the loading corresponds to the metabolite's strength of correlation to that PC. A metabolite with a large magnitude (close to 1 or -1) has a strong influence/contribution to that PC and a metabolite with a small magnitude close to 0 has weak influence/contribution to that PC. A positive loading means that a high value in that metabolite corresponds to a high value in that PC. A negative loading means a high value in that metabolite corresponds to a low value in that PC.  
