---
title: "`r params$title %||% 'Metabolomics Data Quality Report'`"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 2
    toc_float: true
params:
  title:
    value: "Metabolomics Data Quality Report"
  notes:
    value: ""
  plots:
    value: !r character(0)
  p:
    value: !r NULL
  d:
    value: !r NULL
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(dev = "svg")
library(htmltools)
library(dplyr)
library(QCcorrection)
```
# Raw Data
```{r}
QCcorrection:::report_text_data_req()
QCcorrection:::report_text_withheld_columns(params$p, params$d)
```

### Raw Data Summary and Structural Checks

**Raw data summary and structural checks.** 
```{r}
QCcorrection:::report_text_data_inspection()
QCcorrection:::ui_basic_info(params$d$cleaned$df, params$d$cleaned$replacement_counts, params$d$cleaned$non_numeric_cols, params$d$cleaned$duplicate_mets, params$d$cleaned$blank_df, params$d$cleaned$below_blank_threshold)
```

### Filter Missing Values in Raw Data
```{r}
QCcorrection:::report_text_mv_filter()
QCcorrection:::ui_filter_info(params$d$filtered$mv_removed_cols, params$d$filtered$mv_cutoff, params$d$filtered$qc_missing_mets, params$d$filtered$class_metab_all_missing)
```

# Imputation, Signal Drift Correction, and Transformation.

### Missing Value Imputation
```{r}
QCcorrection:::report_text_imputation(params$p, params$d)
```

### Correction Description
```{r}
QCcorrection:::report_text_correction(params$p, params$d)
QCcorrection:::report_text_correction_descriptions()
```

### Post-correction Filtering
#### QC RSD Filtering

```{r}
QCcorrection:::report_text_rsd_cal()
QCcorrection:::ui_postcor_filter_info(params$d$filtered_corrected, params$p$remove_imputed, params$p$rsd_cutoff, params$p$post_cor_filter)
```

#### Candidate Extreme Values

```{r}
QCcorrection:::report_text_ev_detection()
```
```{r, echo=FALSE, fig.width=7.5, fig.height=5.5}
hp <- params$plots[["Hotelling PCA"]]
if (!is.null(hp)) {
  print(hp)
} else {
  htmltools::tags$span(style = "font-weight:bold;", "Hotelling PCA plot was not available.")
}
```
```{r}
QCcorrection:::ui_outliers(params$p, params$d, include_plot = FALSE) 
```

### Transformation Description
```{r}
QCcorrection:::report_text_transformation(params$p, params$d)
QCcorrection:::report_text_transform_methods()
```

###  Metabolite Correlations
```{r}
QCcorrection:::report_text_correlations()
if (!is.null(params$d$all_corr)) {
  QCcorrection:::ui_corr_range_info(params$d$all_corr, params$p$corr_threshold)
} else {
  htmltools::tags$span(style = "font-weight:bold;", "Metabolite correlations were not computed.")
}
```
# Visualization

### Visualize the Correction with Metabolite Scatter Plots
```{r, echo=FALSE, fig.show='hold', out.width='49%', fig.width=5, fig.height=5}
print(params$plots[["Metabolite Scatter 1"]])
print(params$plots[["Metabolite Scatter 2"]])
```
```{r}
QCcorrection:::report_text_scatter_intro(params$p, params$d)
QCcorrection:::report_text_met_scatter()
```

### RSD Comparison
To judge the quality of the cleaned data, we can visualize the change in variation with the relative standard deviation (RSD) comparison plots. RSD is sometimes called coefficient of variation (CV) and is computed as RSD = (standard deviation / mean) * 100%. High RSD indicates that the values for that metabolite are more spread out around the mean, which may be an expected result in the experimental samples. For QC samples, RSD should decrease after correction.
```{r}
QCcorrection:::report_text_rsd_intro(params$p, params$d)
```
```{r, echo=FALSE, fig.width=7.5, fig.height=4.5}
print(params$plots[["RSD Comparison"]])
```
```{r}
QCcorrection:::report_text_rsd_plots()
```

```{r}
QCcorrection:::report_text_rsd_table()
QCcorrection:::ui_rsd_stats("filtered_cor_data", params$p, params$d)
```

### PCA Comparison
```{r}
QCcorrection:::report_text_pca_plots()
```
```{r, echo=FALSE, fig.width=8.333, fig.height=4.417}
print(params$plots[["PCA Comparison"]])
```

```{r, echo=FALSE, fig.width=8.75, fig.height=4.417}
print(params$plots[["PCA Loading"]])
```
